Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-23 paper/Response.tex

\begin{document}

October 29, 2010

Response to reviewers:

We thank the reviewers for their careful reading of our paper.  Their
suggestions helped us improve the clarity of what we discuss, and even
sparked some new ideas for improving our algorithms.  Below we list
the Reviewers' comments; our responses are shown in red.

@@

\vspace{0.5 in}

Reviewer \#1: The submitted paper features a description and analysis of a
new MapReduce implementation based on MPI.  The features and limitations
of the implementation are covered with respect to other MapReduce
implementations.  The implementation is applied to several
graph-theoretic algorithms, and the performance of these applications
is compared with applications from an established library (Trilinos).

The introductory and supporting text are, overall, very good.  The
paper places MapReduce in its historical context and motivates its
usage.  The outline of strengths is helpful- important features
include support for out-of-core operations that other libraries (e.g.,
Trilinos) often cannot handle and the ability to use MPI directly in
addition to the MapReduce abstraction.  The paper then provides a good
overview of the composition of operations in a MapReduce application
and a summary of the out-of-core mechanisms available.  The figure
(Figure 1) used to provide a high-level schematic of MapReduce data
flows could be enhanced- for example, it should indicate node-local
and communication intensive operations.

\revise{Figure 1 is actually describing an on-processor rearrangement
of data that each processor performs on its local data.  Thus there is
no inter-processor communication involved.  We added text to the caption
to clarify that all operations are done on-processor.}

@@

All performance results are shown as scaling studies for a range of
problem sizes for the six graph algorithms.  The results show that the
MapReduce applications do not compete well with the reference
implementation (Trilinos), but the authors make clear the limitations
involved in using this high-level library which trades some
performance for usability.  Additionally, the implementation supports
very large data sets which can only be processed using its out-of-core
feature set.  The results are presented well.

Since improved performance for small and mid-range problems is not the
contribution of the paper, it could be improved by stressing other
aspects of the system.  Real-world out-of-core problems could be
surveyed.  Code size and complexity could be considered with usability
as a motivation.  Is the MapReduce-based graph algorithm library
easier to use than the Trilinos library?  Should higher-level user
applications be considered?  Can the MapReduce-based inefficiencies be
isolated more concisely?

\revise{We added some of these comparison points to a new section 5.7
  and to the final ``Lessons Learned'' section.  In particular we
  highlight the relatively small line count of the MapReduce
  algorithms and the natural out-of-core ability it provides, as
  compared for example to Trilinos.  The comment with respect to MR
  inefficiencies is addressed below.}

@@

\vspace{0.5 in}

Reviewer \#3: I appreciated the clear exposition of your map/reduce
framework, its decomposition into a number of independent and composable
stages, and the explanation of how out-of-core computations are handled.

I also liked that you presented the map/reduce implementation of a
number of different graph algorithms. However, I thought the
presentation of these algorithms could be improved in two ways: - The
text tends to just explaining what the code does, not why, nor why
this solves the specific problem (4.2 was the worst culprit here - f I
understood correctly, the vertices get labeled by edge counts that are
used to give the edges a direction that helps identify triangles: if
this is correct a little diagram showing this would help...)  - The
detailed algorithms (Figures 4-9) are a little too imprecise to follow
at times. In particular, when multiple map/reduces are involved the
lack of names for the map/reduce operations and their inputs makes
things confusing. Please be very explicit (after all the code itself
is), and Figure 4 at least actually had names for the map-reduce
objects.

\revise{We agree that our notation in Section 4 was deficient.  I
  couldn't understand some of the details when I re-read it!  So we
  re-worked and standardized the notation used in Section 4 to
  hopefully describe all the graph algorithms more precisely.  In
  particular we noted which operations are performed with which
  MapReduce objects and their associated key/value data, so that the
  flow of data through each algorithm should be more clear.  A
  discussion of the MR-MPI library's ability to allow creation of
  multiple MapReduce objects was added to Section 2.}

### 2010-11-29 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@677 3fc257bb-8550-0410-81b3-a5c103da83f7
More accurate line counts.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@680 3fc257bb-8550-0410-81b3-a5c103da83f7

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
Algorithm & R-MAT & R-MAT++ & PageRank & Tri & CC & MIS & SSSP & SSSP++ \\
\hline
Lines of code & 208 & 216 & 694 & 367 & 453 & 393 & 505 & 569 \\
Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
\hline
\end{tabular}
\caption{Lines of code to implement each of the graph
  algorithms in the associated figures.}
\label{t:rmat}
\end{center}
\end{table}

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
must implement the parallelism explicitly, whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 1206 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
an entire matrix-vector multiplication can be invoked by a single call to
Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required 697 lines of Java code.

>>>>

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
Algorithm & R-MAT & R-MAT++ & PageRank & Tri & CC & MIS & SSSP & SSSP++ \\
\hline
Lines of code & 208 & 216 & 603 & 367 & 453 & 393 & 394 & 448 \\
Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
\hline
\end{tabular}
\caption{Lines of code to implement each of the graph
  algorithms in the associated figures.}
\label{t:rmat}
\end{center}
\end{table}

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 845 lines of code, mainly because it
must implement the parallelism explicitly, whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 718 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
entire dot-products and matrix-vector multiplications can be invoked by single function calls to
Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required 697 lines of Java.
